Rotation Periods in Low-Mass Binary Stars: The Impact of Tidal Torques and Magnetic Braking

### 2018-12-10 Paper/sync.tex
added some text

The predictions of the CPL model fail to reproduce several key aspects of that data.  First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} \lsim 1.5$, a prediction that is entirely inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data likely owing to the fact that stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.  The CPL model can produce the subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.  The CTL model, however, excels where the CPL models fails and naturally produces both populations, suggesting that the CTL model better describes tidal interactions in low-mass binaries.  We examine this point further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}.

\citet{Claret1995} find that supersynchronous rotation their sample of eccentric binaries are in good agreement with the predictions of tidal models.

@@

Median subsync tidal tau: 0.0790922272
Median supersync tidal tau: 0.12508653609999998
Median subsync ecc: 0.1288805
Median supersync ecc: 0.151426
Fraction with Porb/Prot in [0.92,1.2] for Porb < 10d: 0.6126373626373627
Fraction with Porb/Prot in [0.84,0.92] for Porb < 10d: 0.11401098901098901

- lurie finds that low mass ratio binaries are more likely to be subsync for Porb less than 10 days, makes sense since lower mass secondary means weaker tidal torques.

Although Lurie+2017 attributes differential rotation to the production of the
subsynchronous population, we find that coupled stellar-tidal evolution naturally
produces the population. it's probably a combination of both effects. Compare with Lurie+2017 sample for 2 $<$ Porb $<$ 10 days:
$72\%$ in [0.92,1.2] and $15\%$ in [0.84, 0.92].

@@

@@

>>>>

The predictions of the CPL model fail to reproduce several key aspects of that data.  First, the CPL model cannot reproduce the cluster of supersynchronous rotators with P$_{orb}/$P$_{rot} \lsim 1.1$ for P$_{orb} < 10$ d and instead predicts that all sypersynchronous rotators lie on the line P$_{orb}/$P$_{rot} \lsim 1.5$, a prediction that is entirely inconsistent with the data as no obvious spin-orbit commensurablities, aside from 1:1 synchronization, is present in the data likely owing to the fact that stellar convective envelopes lack a fixed shape, making resonant coupling difficult unless it occurs with internal gravity or pressure modes \citep{Burkart2014,Lurie2017}.  The CPL model can produce the subsynchronous rotators observed for $10 \lsim$ P$_{orb} \lsim 20$ d, but fails to produce the prominent cluster of subsynchronous rotators at P$_{orb}/$P$_{rot} \approx 0.9$ for P$_{orb} < 10$ d observed by \citet{Lurie2017}.  The CTL model, however, seems to excel where the CPL models fails by producing both populations.  We examine these features further by comparing results from our CTL model with the \citet{Lurie2017} data where the data is most complete, P$_{orb} < 10$ d, in $\S$~\ref{sec:subsync}, and offer tests to discriminate between the two models in $\S$~\ref{sec:whichModel}.

@@

Median supersync ecc: 0.15

@@

Although \citet{Lurie2017} argues that differential rotation creates the subsynchronous population, we find that the competition between weak tidal torques and magnetic braking naturally
produces the population.  Our mechanism for producing long-lasting subsynchronous rotation described in $\S$~(XXX sec ref) does produce the population seen in Fig.~\ref{fig:subsync} as the sunsynchronous rotators systematically have weaker tidal torques, with a median $\tau = 0.13$ s, compared to the median $\tau = 0.08$ s for the supersynchronous population in which strong tides tidally-lock the binaries.

Our CTL model successfully reproduces the \citet{Lurie2017} data not just by qualitatively matching the spin-orbital populations observed in \kepler, but also in specific populations. \citet{Lurie2017} find that $15\%$ of their subsynchronous population with 2 $<$ Porb $<$ 10 days has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $11\%$ of our CTL population, suggesting that subsynchronous rotation is a common outcome of stellar-tidal evolution in low-mass binaries. Furthermore in the same P$_{orb}$ regime, \citet{Lurie2017} find that $72\%$ of their population has P$_{orb}/$P$_{rot} \in [0.84, 0.92]$, compared with $61\%$ of our CTL population.   Both are close matches, especially given that we did not hand-tune our simulation initial conditions, suggesting that our coupled stellar-tidal model using the CTL formalism is an accurate description of the physics of tidally-interacting binaries. These findings suggest that the CTL model describes tidal interactions in low-mass binaries better than the CPL model, but in the absence of reliable uncertanties on the \citet{Lurie2017} data, we cannot definitively assert that in general, the CTL model is superior to the CPL model for tidal torques in low-mass binaries.  We discuss means to discriminate between the two models in $\S$~\ref{sec:whichModel}.

@@

\subsection{CPL or CTL?} \label{sec:whichModel}

Here we offer observational tests to discriminate between the two models in short P$_{orb}$ binaries where tidal effects are more likely to be significant.  First, one could obtain a series of time-resolved spectra of the short P$_{orb}$ supersynchronous binaries studied by \citet{Lurie2017} to derive more robust $e$ constraints than derived using the approximations in \citet{Lurie2017}.  For binaries with P$_{orb} < 10$ and $e < \sqrt{1/19}$, the CPL model does not permit supersynchronous rotation except in extremely young systems that are still contracting along the pre-main sequence, or that have just reached the main sequence, while the CTL model does allow for it.  If supersynchronous rotation is observed in such systems, it is evidence in favor of the CTL model.

Second, for tidally-locked binaries with $e > \sqrt{1/19}$ for short P$_{orb}$, the CPL model predicts supersynchronous rotation in the form of a 3:2 spin-orbit comensurability, e.g. the line at P$_{orb}/$P$_{rot} = 1.5$ seen in the left panel of Fig.~\ref{fig:lurie7}, and no other spin state is permited.  The CTL model, however, permits a continuum of supersynchronous states in eccentric tidally-locked rotators, e.g. right panel of Fig.~\ref{fig:lurie7}, that seems to provide a much better match to the \citet{Lurie2017} data, providing strong evidence in favor of the CLT model. The lack of uncertainties on the \citet{Lurie2017} P$_{rot}$ measurements and their poor eccentricity constrains, however, prohibit us from making this distinction conclusive in this work.  Again, more robust $e$ constrains via time-dependent spectra would aid in this discriminaton between models, as would determinations of P$_{rot}$ with uncertainties, perhaps using a gaussian process model as in (XXX cite GP rasumussen, and correct ruth angus paper).  Accurate determinations of the spin-orbital state of short P$_{orb}$ binaries are critical to determine which tidal model best describes the evolution of low-mass stellar binaries.
